\appendix
\section{Complete Experimental Inventory (Template)}
\begin{itemize}[leftmargin=1.2em]
    \item TODO-A1: Master list of all completed runs (run tag, seed, tasks, model, modes).
    \item TODO-A2: Full hyperparameter grids attempted and rejected.
    \item TODO-A3: Cluster configuration and runtime budget.
\end{itemize}

\section{Equation Summary}
\begin{align}
\mathcal{L}&=\lambda_f\mathcal{L}_{\text{fail}}+\lambda_\tau\mathcal{L}_{\tau}+\lambda_c\mathcal{L}_{\Delta p},\\
\mathcal{L}_{\text{fail}}&=\text{BCEWithLogits}(\hat{y}^{\text{fail}}, y; w^+_{\text{batch}}),\\
\mathcal{L}_{\tau}&=\text{Huber}_{\delta}(\hat{\tau},\tau),\\
\mathcal{L}_{\Delta p}&=\text{Huber}_{\delta}(\widehat{\Delta\mathbf{p}},\Delta\mathbf{p}).
\end{align}

\section{Claim-to-Evidence Mapping (Required for Final)}
\begin{itemize}[leftmargin=1.2em]
    \item Claim C1 (transfer detection): TODO-CE-1 result file list.
    \item Claim C2 (stop vs steer): TODO-CE-2 result file list.
    \item Claim C3 (vs MPPI): TODO-CE-3 result file list.
    \item Claim C4 (latency): TODO-CE-4 result file list.
\end{itemize}

\section{Yahboom Embodied Checklist (Required for Final)}
\begin{itemize}[leftmargin=1.2em]
    \item TODO-YBA-1: Exact arm hardware, firmware, and controller stack.
    \item TODO-YBA-2: Camera and calibration details (extrinsics, frame transforms).
    \item TODO-YBA-3: Action interface mapping from policy output to arm commands.
    \item TODO-YBA-4: Safety envelope (joint/EEF bounds, velocity caps, emergency stop).
    \item TODO-YBA-5: Real-robot task list and environment setup photos.
    \item TODO-YBA-6: Dynamic obstacle protocol details (timing, object classes, operator procedure).
    \item TODO-YBA-7: Trial accounting table (per-task N, seeds, resets, failure causes).
    \item TODO-YBA-8: Latency decomposition (policy, safety MLP, controller, bus/actuation).
    \item TODO-YBA-9: Hardware result table (vanilla, latent stop, MPPI, steering).
\end{itemize}

\section{Figures and Tables Checklist}
\begin{itemize}[leftmargin=1.2em]
    \item Figure 1 system architecture (camera-ready art): TODO-FIG-1.
    \item Latent trajectory visualization (PCA/t-SNE): TODO-FIG-2.
    \item Zero-shot OOD final table: TODO-TBL-1.
    \item OpenVLA OOD final table: TODO-TBL-2.
    \item Yahboom hardware final table: TODO-TBL-3.
\end{itemize}

\section{Known Gaps Before Submission}
\begin{itemize}[leftmargin=1.2em]
    \item TODO-G1: Finalize OpenVLA OOD seed-complete results.
    \item TODO-G2: Integrate true OOD BDDL runs (if available) into main table.
    \item TODO-G3: Finalize confidence intervals and statistical tests.
    \item TODO-G4: Finalize Yahboom hardware trial set and include complete embodied analysis.
\end{itemize}
