\section{Discussion}
\label{sec:discussion}

\paragraph{PULSE as inference optimization, not a safety adapter.}
The primary deployable contribution is \textbf{query amortization}: practitioners who want to use a hidden-state failure logit as a closed-loop controller cost need not run $K{=}16$ forward evaluations per step. Single-step gating is TOST-equivalent to $K$-sample search on LIBERO-Spatial at $6.1{\times}$ lower latency. We explicitly do \emph{not} claim certified safe operation, harm reduction, or conformal coverage on control---those require separate calibration (Platt scaling for OpenVLA probabilities; \texttt{platt\_calibrate\_failure\_head.py}) and, for spatial corrections, better supervision than progress-aligned teleportation labels (\texttt{recovery\_conditioned\_labels.py}).

\paragraph{Spatial--control decoupling.}
The \emph{spatial--control decoupling} in \S\ref{sec:mechanism} is a methodological lesson: offline directional AUC on time-matched labels must not be read as closed-loop steering performance---latent jiggle preserves magnitude gating and matches learned direction on episode success. On LIBERO-Spatial, the failure head alone suffices for the latency result; the correction head is a diagnostic under the current label regime.

\paragraph{Correction labels and calibration.}
The directionality analysis (Appendix~\ref{app:spatial_failure}) shows hidden states encode failure geometry early in rollouts, but teleportation labels $\Delta\mathbf{p}_t = \mathbf{p}^{\mathrm{succ}}_t - \mathbf{p}^{\mathrm{fail}}_t$ at matched progress fractions do not translate into control gain. \textbf{Recovery-conditioned matching} (monotonic forward nearest point on a success reference path) replaces that supervision; retraining with \texttt{--label-mode recovery} is the intended path to re-couple spatial readout to control. OpenVLA's raw sigmoid is overconfident (ECE $0.22$); affine Platt scaling on logits is monotone and preserves AUC while collapsing calibration error on the held-out test fold.

\paragraph{Limitations.}
All quantitative controller results are on LIBERO-Spatial. Cross-architecture linear transfer fails. SO-101 hardware validates whether sim-trained probes and the latency advantage survive real control loops and variable-length failures. Head-to-head comparison with FIPER and FAIL-Detect on our rollout pool remains open.
