\section{Experimental Setup}

\subsection{Benchmarks and Task Suites}
Primary evaluation uses LIBERO spatial tasks with off-screen rendering and fixed initialization protocols.

\subsection{Models}
\begin{itemize}[leftmargin=1.2em]
    \item OpenVLA-7B OFT (foundation policy)
    \item ACT (imitation policy)
    \item Auxiliary collected evidence: Diffusion Policy, Octo (where available)
\end{itemize}

\subsection{Evaluation Modes}
\begin{itemize}[leftmargin=1.2em]
    \item vanilla
    \item latent\_stop
    \item noise
    \item ema\_only
    \item latent\_jiggle
    \item mppi
    \item steering (ours)
\end{itemize}

\subsection{Generalization Protocols}
\paragraph{Strict zero-shot task split.}
Train safety head on task subset A, evaluate on disjoint unseen subset B.

\paragraph{OOD stress testing.}
Use disturbance OOD and/or custom OOD BDDL environments.

\subsection{Embodied Platform Track: Yahboom Arm}
This project includes an embodied track on a Yahboom 7-DoF arm controlled through a Jetson-hosted HTTP API.

\paragraph{Current hardware/software stack.}
\begin{itemize}[leftmargin=1.2em]
    \item \textbf{Robot:} Yahboom 7-DoF arm + gripper, serial link on \texttt{/dev/ttyUSB0}.
    \item \textbf{Edge compute:} Jetson Orin running \texttt{arm\_server.py} with camera stream and action API.
    \item \textbf{Control API:} \texttt{/status}, \texttt{/snapshot}, and \texttt{/action} endpoints for \texttt{move\_to}, \texttt{send\_angles}, \texttt{set\_gripper}, \texttt{pick\_at}, \texttt{place\_at}.
    \item \textbf{Host bridge:} \texttt{scripts/yahboom\_bridge.py} and \texttt{scripts/run\_model\_yahboom\_loop.py}.
\end{itemize}

\paragraph{Camera and policy-input standardization.}
The live Jetson camera stream is consumed via \texttt{/snapshot} at 640$\times$480. Before policy inference, images are center-cropped and resized to 224$\times$224 to match the LIBERO-style square-input regime used in simulation pipelines. We additionally run a camera diagnostic script (\texttt{scripts/check\_jetson\_camera\_quality.py}) that records brightness, contrast, sharpness (Laplacian variance), and a center-region occlusion proxy in canonical arm poses (home/watch).

\paragraph{Embodied inference details.}
OpenVLA is loaded with \texttt{trust\_remote\_code=True}. On multi-GPU hosts, sharded loading (\texttt{device\_map=auto}) is used to avoid single-GPU OOM. A safety interlock suppresses motion deltas when action outputs are saturated, preventing unsafe command bursts.

\paragraph{Real-robot protocol status.}
Bring-up and closed-loop command validation are complete (API, gripper, and bounded motion). Standardized autonomous pick/place success-rate benchmarking remains pending and is reported explicitly as future work.

\subsection{Metrics}
Success rate, $\Delta$ vs vanilla, $\Delta$(Ours$-$MPPI), intervention rate, correction magnitude, and controller apply latency (mean/p95).

\subsection{Statistical Reporting Plan for Camera-Ready}
TODO-STATS-1: number of seeds. \\
TODO-STATS-2: confidence interval method. \\
TODO-STATS-3: hypothesis tests. \\
TODO-STATS-4: model-comparison correction.
