\section{Introduction}

Robot foundation models---vision-language-action policies and imitation-learning architectures~\citep{openvla_paper, act_paper}---are increasingly deployed at control rates where every millisecond of extra inference matters. When a probe on hidden states is used as a \emph{controller cost}, planners that query it $K$ times per step multiply backbone-adjacent latency. We treat \textbf{PULSE} as a \textbf{low-latency inference optimization} layer: a frozen-policy add-on that exposes a failure logit for single-query gating, not as a certified safety filter (no dynamics model, no conformal control guarantee, no shield).

\paragraph{Scope (inference vs.\ safety).}
Throughout, \emph{safety} means an empirical failure-anticipation signal usable as a controller cost. PULSE does not provide formal safe-set invariance or calibrated abstention unless post-hoc calibration is applied (Platt scaling reduces OpenVLA ECE from $0.22$ on raw sigmoids; see \texttt{scripts/platt\_calibrate\_failure\_head.py}). Headline claims are \textbf{detection AUC} and \textbf{query amortization} (single-step $\equiv$ $K$-sample search on success), not guaranteed harm reduction.

A natural question is whether the failure signal already exists inside the policy's own representations. SAFE~\citep{safe_paper} showed that a linear probe on VLA hidden states can predict binary failure. We ask three follow-on questions: (a) does a nonlinear probe extract more failure-relevant signal than a linear one on the same activations, and does the same probe template apply across architecturally distinct VLAs without per-architecture tuning; (b) is $K$-sample search over the failure head necessary to use it as a controller cost, or does a single query per step suffice; (c) under standard time-index-matched correction labels, does the learned correction direction add to detection timing in the resulting controller. We train \textbf{PULSE} on frozen hidden states from OpenVLA-7B and ACT, and answer: yes (C1, C2), no (C3), and no (C4).

\paragraph{Contributions.}
\begin{itemize}[leftmargin=1.2em, topsep=2pt, itemsep=2pt]
    \item[\textbf{C1}] \textbf{One probe template, two architectures, no per-architecture tuning.} The same MLP --- input LayerNorm $\to$ $256 \to 128 \to 64$ bottleneck $\to$ three linear heads --- trains from scratch on OpenVLA-7B hidden states (4096-d, autoregressive transformer) and ACT hidden states (256-d, CVAE encoder), a $16\times$ feature-dimensionality gap, with identical hyperparameters and training recipe.
    \item[\textbf{C2}] \textbf{Nonlinear probe gains $+0.05/+0.10$ failure-detection AUC over a linear probe on the same hidden states.} On the trajectory-disjoint test split, the failure head reaches AUC $0.832$ (OpenVLA) and $0.889$ (ACT) vs.\ $0.785/0.788$ for L2-regularised logistic regression on the same activations. The gap isolates the contribution of the nonlinearity: failure is linearly readable, but a small MLP extracts measurably more.
    \item[\textbf{C3}] \textbf{$K{=}1$ query to the failure head is sufficient; $K{=}16$ search adds no success.} Two controllers differ only in how many times the failure head is queried per step. $K{=}16$ MPPI scores $16$ candidate corrections per step and picks the minimum-failure-probability one ($5.6$\,ms). Single-step gating queries the head once and applies a fixed correction on a positive ($0.9$\,ms). Pooled success is TOST-equivalent within ${\pm}2$\,pp ($p{=}0.011$) at $6.1\times$ lower per-step latency. Per-task intervention rate further correlates with vanilla task success at $r{=}{-}0.981$ ($p{=}5.2{\times}10^{-7}$, $n{=}10$) --- the gate fires densely on hard tasks and sparsely on easy ones with no explicit difficulty signal. The claim is about query efficiency on this head, not superiority to other safety controllers.
    \item[\textbf{C4}] \textbf{Spatial--control decoupling under time-matched labels.} Latent jiggle (random direction, matched MLP magnitude gate) matches learned steering on pooled success ($52.90\%$ vs.\ $51.74\%$; TOST $p{=}0.119$ at $\pm 2$\,pp). High offline directional AUC (Table~\ref{tab:safety_head_metrics}) does not imply closed-loop steering gain (\S\ref{sec:mechanism}): timing and magnitude carry the payload under micro-scale adjustments on LIBERO-Spatial. The latent still encodes early failure direction (Appendix~\ref{app:spatial_failure}); re-coupling spatial readout to control requires recovery-conditioned labels or earlier gating.
\end{itemize}
